
\aaa{The main theorem of diagonalization}

\[thm]{
The following conditions for $A$ are equivalent
\[itemize]{
\item $A$ is diagonalizable
\item $A$ satisfies a polynomial of simple roots
\item $A$ is a linear combination of projection matrices $P_1,...,P_k$ such that $P_1+\cdots+P_k=I$.
}
}

We prove this theorem by proving 

¡ “Condition 1“ ⟹  “Condition 2“ ⟹  “Condition 3“ ⟹  “Condition 1“

\a\aa
Condition 1 ⟹   Condition 2.

Suppose $A$ is diagonalizable, we can write
$$
A = P^{-1} \m{λ_1}{}{}{}{},{}{λ_2}{}{}{},{}{}\ddots{}{},{}{}{}{λ_{n-1}}{},{}{}{}{}{λ_n}. P.
$$
Note that some of the diagonal entries might be repeated, therefore, we collect them into a set
$$
S = \{λ_1\} ∪ \{λ_2\} ∪ \cdots ∪ \{λ_n\}
$$
therefore we compose polynomial
$$
f(t)=∏_{μ_i ∈ S}(t-μ_i)
$$
\a\aa
We have
$$
f(A) = P^{-1} \m{f(λ_1)}{}{}{}{},{}{f(λ_2)}{}{}{},{}{}\ddots{}{},{}{}{}{f(λ_{n-1})}{},{}{}{}{}{λ_n}. P=0
$$
\a\aa
Condition 2 ⟹  Condition 3.

Now if $A$ satisfies 
$$
(A-λ_1)(A-λ_2)\cdots(A-λ_n)=0
$$
for \x{distinct} $λ_1,\cdots,λ_n$, 
then we may construct interpolation polynomials $f_{λ_1},\cdots,f_{λ_n}$ with the property that $f_{λ_j}(λ_i)=0$ for any $i≠j$ and $f_{λ_i}(λ_i)=1$. Using interpolation theorem, we interprating the constant function $g(t)=1$, we have
$$ 1 = Q(t)(t- λ_1)(t-λ_2)\cdots(t-λ_n) + f_{λ_1}(t) + f_{λ_2}(t)+\cdots+f_{λ_k}(t) $$
letting $g(t)=t$, we have
$$ t = Q(t)(t- λ_1)(t-λ_2)\cdots(t-λ_n) +\x{λ_1} f_{λ_1}(t) + \x{λ_2}f_{λ_2}(t)+\cdots+\x{λ_k}f_{λ_k}(t) $$
\a\aa
Plug in $t=A$ to those two polynomials, we denote
$$
P_{λ_i}:= f_{λ_i}(A),
$$
we know $P_{λ_i}$ are projection matrices, therefore
$$
I = P_{λ_1}+P_{λ_2}+\cdots+P_{λ_k}
$$
and
$$
A =  λ_1P_{λ_1}+λ_2P_{λ_2}+\cdots+λ_kP_{λ_k}
$$

\a\aa
Condition 3 ⟹  Condition 1.

Recall that by cross filling, we essentially decompose each projection matrices into rank 1 projection matrices

$$
P = \vec v_1\vec w_1^T +\cdots 
$$

Therefore, $A$ is also a linear combination of rank-1 proejction matrices
$$ A = μ_1\vec v_1\vec w_1^T + μ_2\vec v_2\vec w_2^T +\cdots+μ_n\vec v_n\vec w_n^T.  $$
$$ I = \vec v_1\vec w_1^T + \vec v_2\vec w_2^T +\cdots+\vec v_n\vec w_n^T.  $$
Note that each rank one projection must be trace one. Since the trace of $I$ is $n$, we know there are \x{exactly} n many rank-1 projections.
Therefore we may write it into matrix form
$$
A = 
\m
||\cdots|,
{\vec v_1}{\vec v_2}\cdots{\vec v_n},
||\cdots|.
\m{μ_1}{}{},{}\ddots{},{}{}{μ_n}.
\m
-{\vec w_1^T}-,
%-{\vec w_2^T}-,
\vdots\vdots\vdots,
-{\vec w_n^T}-.
$$
\a\aa
and
$$
\m
||\cdots|,
{\vec v_1}{\vec v_2}\cdots{\vec v_n},
||\cdots|.
\m
-{\vec w_1^T}-,
%-{\vec w_2^T}-,
\vdots\vdots\vdots,
-{\vec w_n^T}-.
=I$$
Since the two matrices are square matrices, so let 
$$
P = 
\m
||\cdots|,
{\vec v_1}{\vec v_2}\cdots{\vec v_n},
||\cdots|.
$$
Then
$$P^{-1}
=
\m
-{\vec w_1^T}-,
%-{\vec w_2^T}-,
\vdots\vdots\vdots,
-{\vec w_n^T}-.
$$
We have $A=P
\m{μ_1}{}{},{}\ddots{},{}{}{μ_n}.
P^{-1}$, therefore $A$ is diagonalizable.


%% Spectural decomposition of orthogonal reflection.

\a\aa
Let's think about the following equivalent conditions

\[itemize]{
\item $A$ is diagonalizable 
\item $A$ satisfies a polynomial of simple roots
\item $A$ is a linear combination of projection matrices $P_1,...,P_k$ such that $P_1+\cdots+P_k=I$.
}

\exe Suppose $A=λ_1P_1+λ_2P_2+\cdots+λ_mP_m$ is a linear combination of matrices such that $P_1+P_2+\cdots+P_m$ is a projection, is $A$ diagonalizable? why?

\a\aa
\sol Denote by $P := P_1+P_2+\cdots+P_m$, since $P$ is a projection, so is $I-P$. Let $P_{m+1}=I-P$, we have
$$
P_1+P_2+\cdots+P_m+P_{m+1}=I
$$
and
$$A=λ_1P_1+λ_2P_2+\cdots+λ_mP_m+\underbrace0_{λ_{m+1}}P_{m+1}$$
Therefore $A$ is diagonalizable

\a\aa

Can you find a linear combination of projection matrices such that it is not diagonalizable?
\a\aa
$$
\m11,00.+\m00,01. = \underbrace{\m11,01.}_{“NOT DIAGONALIZABLE!“}
$$


\a\aa
\exe Diagonalize the following matrix
$$
\m211,121,112.
$$

\a\aa
\exe If
$$
A = λ_1P_1+\cdots+λ_mP_m
$$
however, 
$$P_1+\cdots+P_m≠I,$$
 but 
$$(P_1+\cdots+P_m)^2=P_1+\cdots+P_m$$
is a projection. Is $A$ a diagonalizable matrix ? why?
\aaa
